The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}

### 2017-10-04 GRB111117A.tex
Addressing referee

The host lies in the brighter end of the expected sGRB host brightness distribution
at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
redshift determination, we find that between 43 and 71 per cent of all sGRB redshifts
should be missed due to host faintness for hosts at $z\sim2$. The high redshift of
GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
the predicted redshift distribution of sGRBs, which is very sensitive to
high-$z$ sGRBs.

@@

There is mounting evidence that short-duration $\gamma$-ray bursts
come from the merger of NSs, either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The classification of GRBs in two groups, initially comes from the bimodal
distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
where the duration T$_{90} < 2$ has been regarded as the dividing line between long
and short GRBs. Additionally, it has been found for lGRBs that there is a
spectral lag in the arrival-time of photons, with the most energetic ones
arriving first. This lag is consistent with zero for sGRBs
\citep{Norris2006}. Because both populations have continuous, overlapping
distributions in their observables and because telescopes observe in differing
bands, it is difficult to impose a single demarcation criterion between the two
classes. For this reason, the distinction between long and short GRBs is
preferably based on a combination of high-energy properties \citep{Zhang2009,
	Bromberg2012a, Bromberg2013}.

The \textit{Swift} satellite \citep{Gehrels2004} greatly improved the
understanding of sGRB progenitors thanks to its quick localization capability.
The bulk of these localizations have associated galaxies at relatively low
redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, and because
most of these measurements come from the associated hosts, it is arguably biased
towards lower redshifts. The host galaxies of sGRBs are diverse. They are more
massive and less actively star-forming on average than lGRB hosts
\citep{Fong2013b}, while in some cases, no host galaxy can be identified above
detection threshold of deep follow-up observations \citep{Berger2010a,
	Tunnicliffe2014}. Together with their position within their hosts
\citep{Fong2013a}, this suggests a progenitor system that can be very long lived
in comparison to lGRBs, and is tracing with host stellar mass rather than
star-formation rate (SFR).

@@

\begin{figure*}
	\centering
	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
	\caption{SED fit showning the best-fist SED to the derived photometry. The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes. Slit losses are responsible for discrepancies between the measured photometry and the X-shooter spectrum. The reason for the spectral gaps at 5500 \AA{} and 10000 \AA{} is from the merging of the arms.}
	\label{fig:SED}
\end{figure*}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run in physical
mode. Because the echelle orders are curved across each detector, a
rectification algorithm is employed which introduces correlations between
neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
maximal resolution. The observations are combined and extracted using scripts
described in Selsing et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
where the full spectral point spread function is modeled across each arm and
used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
slit position is shown in Fig.~\ref{fig:spec_setup}. We show the extracted
spectrum in Fig.~\ref{fig:SED}.

@@

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$5007,
and \ha. \hb{} is detected at low significance ($\sim$ 3-$\sigma$). We show
\oiii$\lambda$5007 in the insert in Fig.~\ref{fig:line}.  \ha{} is only visible in
the first epoch due to the $K$-band blocking filter used for the remaining
observations. The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and
show significant velocity structure along the slit. A drop in the continuum
bluewards of the \lya{} line further supports the inferred redshift.

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed at
21\,000~\AA{} where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. Due to their velocity structure, the lines exhibit clear
deviations from a Gaussian and given the low S/N of the spectra we do not attempt
any parametric fits. We instead obtain a limit on the SFR by numerically
integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} > 4.1
\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After converting
the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass
function using \citet{Madau2014}, we derive a limit of $SFR > 7
M_\odot$~yr$^{-1}$. 
%Additionally, based on the width of \oiii, an aperture
%covering \ha is integrated over, where sky lines have been masked and
%interpolated over, using a synthetic sky spectrum \citep{Noll2012, Jones2013}.
%From the integrated \ha-line, we estimate SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.
%The error on the SFR is derived from the integrated variance spectrum over the
%\ha-line and excluding the intrinsic scatter of the relation. 
From the SED-fit
(Sect.~\ref{SED}), and the detection of \lya, the host is constrained to contain
very little or no dust, although the presence of Lya does not exclude dust.
Therefore we do not apply a dust-correction to the measured \ha{} flux here.
\oii{} is close to a region of strong telluric absorption, which is why no SFR is
inferred from this line.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Most data were taken long after the GRB when no
afterglow contribution was present. Given the faintness of the afterglow (see
Sect. \ref{xray}, and \citealt{Cucchiara2011, Cenko2011}), we also expect
negligible contribution to the earliest epochs, which is confirmed by the
consistency between the two $g$-band measurements.

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (where
necessary), sky-subtraction, and stacking of individual images. On the final
reduced products, seeing-matched aperture photometry was carried out using
\texttt{Source Extractor} \citep{Bertin1996} where the size of the aperture was
chosen to be twice the seeing FWHM. (THOMAS! WAS THE APERTURE FIXED BETWEEN
FILTERS AND IS THIS CORRECT?) Photometric calibration was fixed relative to
field stars from the SDSS and 2MASS catalogs in the case of $griz$ and
$JHK_{\mathrm{s}}$ filters, respectively. For the $R$- and $I$-band photometry,
we used the color transformations of Lupton%
\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We
convert all magnitudes into the AB system, and correct for a Galactic foreground
of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.

Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the
re-analysis of the Gemini data compared to the $z$-band measurements of
\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
$z$-band magnitude that is brighter by 0.8-1.0 mag compared to our value, where
data taken in other filters are consistent within the errors. Visual inspection
of the Gemini image shows only a marginal detection of the host, which we report
here as a $\sim$ 2-$\sigma$ measurement. More conservatively, the 3-$\sigma$
upper limit for the Gemini image is $z > 24.06$. Objects of magnitude $z \sim
23$ are clearly seen and are significantly brighter than the GRB host galaxy.
The consistency between the deeper FORS2 $z$-band image and the upper limit
derived for the Gemini image lends credence to the inferred magnitudes presented
here, see Fig. \ref{photplot}.

@@

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
afterglow limits \citep{Cucchiara2011, Cenko2011}, the X-ray lightcurve can be
interpolated and evaluated at the time of the non-detection. We find $\beta_{OX}
< 0.79$, consistent with what was reported in \citet{Sakamoto2013}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of the restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As pointed out by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is
securely classified as a sGRB. Because the observed classification indicators,
$T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation.
The intrinsic spectral lag shortens, but since it is already consistent with
zero, this does not affect the classification.

The intrinsic luminosity is shown in the X-ray light curve (Fig.
\ref{fig:sxray_lightcurve}) and it is sub-luminous compared to the majority of
long GRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
distribution at 10 ks. The sub-samples comprise of 333 long, 19 short GRBs, and
GRB~111117A. The mean and the 1-$\sigma$ dispersions of the samples are
L$_{lGRB} = 46.59 \pm 0.87$ and L$_{sGRB} = 44.96 \pm 0.94$. GRB~111117A had a
luminosity of 44.95 at 10 ks which is very close to the peak of the sGRB
luminosity distribution at 10 ks, but an outlier from the lGRG distribution,
further supporting the short classification.

\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 \citep{DeUgartePostigo2006, Levan2006} at $1.7
\lesssim z \lesssim 4.5$ ($17_{-15}^{+14}$ per cent) and GRB~090426
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.

Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
intermediate duration bursts. This comes from both events having very soft
spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
bursts are very clearly related in their properties to lGRBs
\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
mergers.

@@

The large offset of GRB~111117A relative to the host center derived in
\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
galaxy-scale gas providing the X-ray absorption. Along with the low dust content
of the host, the large offset from the host center indicates that the high $N_H$
arises because the density in the GRB surroundings is high, or because the light
from the afterglow traverses a region of dense gas. Assuming a dust-to-metals
ratio typical for the Milky Way, the X-ray derived $N_H$ corresponds to significant
extinction along the afterglow sight line ($A_V > 10$) \citep{Predehl1995},
which is contrasted to the absence of dust found for the SED fit. The discrepancy between the extinction derived from the GRB afterglow and the one obtained using galaxy-wide measures is also seen in the range of ratios found for lGRBs \citep{Perley2013a, Covino2013}.

@@

The detection of \lya{} is consistent with the SED-inferred absence of dust,
despite the moderate stellar age which would suggest the opposite. The centroid
of the Ly$\alpha$ emission is found to be redshifted by $\sim$~220~km~s$^{-1}$
with respect to systemic, which is similar to what is found for long GRB hosts
\citep{Milvang-Jensen2012a} where the outflow is attributed to star formation.

@@

@@

Using the complete, flux-limited selection of bursts from \citet{DAvanzo2014a},
excluding GRB~111117A and the likely non-sGRB GRB~090426, we have a
statistically homogeneous sample from which we can address the implications of
the redshift of GRB~111117A. Out of the 14 hosts in the sample, 10 (71 per cent)
have measured redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the
host of GRB~111117A  is brighter than 80 per cent of the hosts with measured
$L_B/L^{\star}_{B}$. Even if we conservatively assume that \textit{all} the
hosts missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the
host is still brighter than > 60 per cent of sGRB hosts. For all 26 hosts with
$L_B/L^{\star}_{B}$ from \citet{Berger2014}, the host of GRB~111117A is brighter
than 73 per cent.

@@

A fraction of the bursts missing redshift are host-less but appear to be
spatially correlated with galaxies that are likely at moderate redshifts
\citep{Tunnicliffe2014}, but should some of the remainder be at high redshift,
the missed fraction will increase. If we assume that \textit{all} the bursts
that are missing redshifts are at high-$z$ and missed due to host faintness, 10
out of 14 hosts in the complete sample (71 per cent) would be missed at $z =
2.211$. This serves as an upper limit on the fraction of missed burst redshifts
at high-$z$. Conversely, if all bursts missing redshift are at low redshift and
missed for other reasons, 6 out of 14 hosts (43 per cent) would be missed at $z
= 2.211$. The two limits indicate that we would miss between 43 and 71 per cent
of sGRB hosts at $z \sim 2$ due to host faintness.

>>>>

The host lies in the brighter end of the expected sGRB host brightness
distribution at $z = 2.211$, and is actively forming stars. Using the host as a
benchmark for redshift determination, we find that between 43 and 71 per cent of
all \textit{Swift}-detected sGRB should be missing redshifts due to host
faintness for bursts at $z\sim2$. The high redshift of GRB~111117A is evidence
against a lognormal delay-time model for sGRBs through the predicted redshift
distribution of sGRBs, which is very sensitive to high-$z$ sGRBs.

@@

There is mounting evidence that short-duration $\gamma$-ray bursts (sGRBs)
come from the merger of neutron stars (NSs), either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Berger2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b, Kann2011}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts, \citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The classification of GRBs in two groups, initially comes from the bimodal
distribution of burst duration and spectral hardness \citep{Kouveliotou1993},
where the duration T$_{90} < 2$ sec has been regarded as the dividing line between long
and short GRBs. Additionally, it has been found for lGRBs that there is a
spectral lag in the arrival-time of photons, with the most energetic ones
arriving first. This lag is consistent with zero for sGRBs
\citep{Norris2006}. Because both populations have continuous, overlapping
distributions in their observables and because telescopes observe in differing
bands, it is difficult to impose a single demarcation criterion between the two
classes. For this reason, the distinction between long and short GRBs is
preferably based on a combination of high-energy properties \citep{Zhang2009,
	Bromberg2012a, Bromberg2013}.

The \textit{Swift} satellite \citep{Gehrels2004} greatly improved the
understanding of sGRB progenitors thanks to its quick localization capability.
The bulk of these localizations have associated galaxies at relatively low
redshifts with a median redshift of $z\sim0.5$ \citep{Berger2014}, and because
most of these measurements come from the associated hosts, it is potentially
biased towards lower redshifts due to Malmquist bias. The host galaxies of sGRBs
are diverse. They are more massive and less actively star-forming on average
than lGRB hosts \citep{Fong2013b}, while in some cases, no host galaxy can be
identified above detection threshold of deep follow-up observations
\citep{Berger2010a, Tunnicliffe2014}. Together with their position within their
hosts \citep{Fong2013a}, this suggests a progenitor system that can be very long
lived in comparison to lGRBs, with the host stellar mass affecting the sGRB
rates more than star-formation rate (SFR) \citep{Berger2014}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=16cm]{figures/SEDspecphot.pdf}
	\caption{SED fit showing the best-fit SED to the derived photometry. The detection of \lya{}  is predicted from the SED fit and confirmed by the spectroscopic observations. Overplotted in grey is the observed spectrum, binned by 6 \AA{} for presentation purposes. An attempt to correct for slit losses has been carried out based on the average seeing of the observations. The blue, dashed line is the corresponding error spectrum, smoothed for presentation purposes. The reason for the spectral gaps at 5500 \AA{} and 10000 \AA{} is from the merging of the arms.}
	\label{fig:SED}
\end{figure*}

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). Observations use a simple ABBA nodding pattern, using 5
arcsec nod throws. X-shooter covers the wavelength range from 3000~\AA{} to
24\,800~\AA{} (21\,000~\AA{} when the $K$-band blocking filter is used) across
three spectroscopic arms. We carried out the bias-correction, flat-fielding,
order tracing, wavelength calibration, rectification, and flux calibration using
the VLT/X-shooter pipeline version 2.8.4 \citep{Goldoni2006, Modigliani2010} run
in physical mode. Because the echelle orders are curved across each detector, a
rectification algorithm is employed which introduces correlations between
neighboring pixels. We select a pixel-scale of 0.2/0.2/0.6 \AA/pix for the
UVB/VIS/NIR arm to minimize the degree of correlation while conserving the
maximal resolution. The observations are combined and extracted using scripts
described in Selsing et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}},
where the full spectral point spread function is modeled across each arm and
used for the optimal extraction algorithm \citep{Horne1986}. An overview of the
spectroscopic observations is given in Table~\ref{tab:spec_overview}, and the
slit position is shown in Fig.~\ref{fig:spec_setup}. Each of the epochs are
extracted individually and combined in a weighted fashion where the weights at
each pixel is chosen as median variance spectrum of the region surrounding that
pixel, thus avoiding to base the weight on the pixel variance. An attempt to
correct for slit-losses has been carried out on the combined spectrum based on
the average seeing of the observations. We show the extracted spectrum in
Fig.~\ref{fig:SED}.

@@

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines belonging to \lya, \oii$\lambda$3727, \hb, \oiii$\lambda$4959,
\oiii$\lambda$5007, and \ha. \hb{} and \oiii$\lambda$4959 are both detected at
low significance ($\sim$ 3-$\sigma$). We show \oiii$\lambda$5007 in the insert
in Fig.~\ref{fig:line}.  \ha{} is only visible in the first epoch due to the
$K$-band blocking filter used for the remaining observations. The nebular lines
exhibit a spatial extent of $\sim$ 1\farc5 and show significant velocity
structure along the slit. A drop in the continuum bluewards of the \lya{} line
further supports the inferred redshift. No spectral evolution is observed across
the epochs, supporting the lack of afterglow contribution.

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of the
host \citep{Kennicutt1998}. At the redshift of the GRB host, \ha{} is observed
at 21\,000~\AA{} where the night sky is very bright. In addition, several bright
sky-lines are superposed on the line, making an accurate estimate of the
\ha-flux difficult. Due to their velocity structure, the lines exhibit clear
deviations from a Gaussian and given the low S/N of the spectra we do not
attempt any parametric fits. We instead obtain a limit on the SFR by numerically
integrating the part of \ha{} free of contamination and obtain $F_{H \alpha} >
4.1 \times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$. After
converting the \citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial
mass function using \citet{Madau2014}, we derive a limit of $SFR > 7
M_\odot$~yr$^{-1}$. We additionally obtain the \lya{} line flux by numerically integrating
the entire \lya{} line complex and obtain $F_{\mathrm{Ly}\alpha} = 2.0 \pm 0.5
\times 10^{-17} \mathrm{erg} \mathrm{s}^{-1} \mathrm{cm}^{-2}$.

%Additionally, based on the width of \oiii, an aperture
%covering \ha is integrated over, where sky lines have been masked and
%interpolated over, using a synthetic sky spectrum \citep{Noll2012, Jones2013}.
%From the integrated \ha-line, we estimate SFR = $18 \pm 3$ M$_\odot$ yr$^{-1}$.
%The error on the SFR is derived from the integrated variance spectrum over the
%\ha-line and excluding the intrinsic scatter of the relation.

From the SED-fit (Sect.~\ref{SED}), and the detection of \lya{}, the host is
constrained to contain very little or no dust, although the presence of \lya{}
does not exclude dust. Therefore we do not apply a dust-correction to the
measured \ha{} flux here. \oii{} is close to a region of strong telluric
absorption, which is why no SFR is inferred from this line.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}. Most data were taken long after the GRB had faded
when no afterglow contribution was present. Given the faintness of the afterglow
(see Sect. \ref{xray}, and \citealt{Cucchiara2011, Cenko2011}), we also expect
negligible contribution to the earliest epochs, which is confirmed by the
consistency between the two $g$-band measurements.

All data were reduced, analyzed and fitted in a similar manner as described in
detail in \citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}.
Briefly, we use our own \texttt{Python} and IRAF routines to perform a standard
reduction which includes bias/flat-field correction, de-fringing (where
necessary), sky-subtraction, and stacking of individual images. On the final
reduced image products, \texttt{DAOPHOT} \citep{Stetson1987} was used to derive
the reported photometry, where the size of the aperture was chosen to be
2\farc0. Because image quality ranges from 0\farc6 - 1\farc7 and the galaxy
major axis is $\sim$ 1\farc, this ensures that, in all cases, the large majority
of the flux is collected and low-surface brightness light missed in the aperture
will not influence the measurement. This method sacrifices some S/N in the best
seeing cases for more reliable photometry across differing observing conditions.

Photometric calibration was fixed relative to field stars from the SDSS and
2MASS catalogs in the case of $griz$ and $JHK_{\mathrm{s}}$ filters,
respectively. For the $R$- and $I$-band photometry, we used the color
transformations of Lupton%
\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. We
convert all magnitudes into the AB system, and correct for a Galactic foreground
of $E_{B-V}=0.027~\mathrm{mag}$ \citep{Schlegel1998, Schlafly2011}.

Noteworthy is the discrepancy of both our new VLT/FORS2 photometry and the
re-analysis of the Gemini data compared to the $z$-band measurements of
\citet{Margutti2012} and \citet{Sakamoto2013}. Both of these authors report
$z$-band magnitudes that are brighter by $\sim 1.0$ mag compared to our values,
where data taken in other filters are consistent within the errors. Visual
inspection of the Gemini image shows only a marginal detection of the host,
which we report here as a $\sim$ 2-$\sigma$ measurement. More conservatively,
the 3-$\sigma$ upper limit for the Gemini image is $z > 24.06$. Objects of
magnitude $z \sim 23$ are clearly seen and are significantly brighter than the
GRB host galaxy. The consistency between the deeper FORS2 $z$-band image and the
upper limit derived for the Gemini image lends credence to the inferred
magnitudes presented here, see Fig. \ref{fig:photplot}.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{figures/photometry_mosaic.pdf}
	\caption{Mosaic showing all obtained imaging. The host of GRB~111117A is marked by a red circle with a 2\farc0 radius. This is the same size as the aperture used for deriving the photometry. Worth noting is the relative depth of the GMOS-N $z$-band and the FORS2 $z$-band. For reference, the object located to the south-east of the host and marked with a blue, dashed circle has an extinction corrected magnitude of $23.11 \pm 0.09$ ($23.10 \pm 0.18$) in the FORS2 (GMOS-N) $z$-band.}
	\label{fig:photplot}
\end{figure}

@@

A measure of the optical-to-X-ray flux ratio is parametrized in terms of the
"darkness"-parameter $\beta_{OX} $ \citep{Jakobsson2004}. Using the optical
afterglow limits ($r' > 25.5$, 13.5 hrs post-burst) \citep{Cucchiara2011,
	Cenko2011}, the X-ray lightcurve can be interpolated and evaluated at the time
of the non-detection. We find $\beta_{OX} < 0.79$, consistent with what was
reported in \citet{Sakamoto2013}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of the restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Based on the BAT light-curve, $T_{90}=0.46$~s which is shorter than both the
prototypical 2 s \citep{Kouveliotou1993} and the $\sim$0.7 s suggested to also
exclude the shorter tail of the \textit{Swift}-observed lGRB population
\citep{Bromberg2012a}. Additionally, no signs of extended emission was found by
\citep{Sakamoto2013}. A spectral lag of $0.6 \pm2.4$ ms, consistent with zero,
is not uniquely associated to sGRBs \citep{Bernardini2014}, but in conjunction
with the duration and the spectral hardness \citep{Sakamoto2011}, as pointed out
by \citet{Margutti2012} and \citet{Sakamoto2013}, GRB~111117A is securely
classified as a sGRB. Because the observed classification indicators, $T_{90}$
and hardness ratio, do not depend strongly on redshift \citep{Littlejohns2013a},
the updated redshift does not change this designation. The intrinsic spectral
lag shortens, but since it is already consistent with zero, this does not affect
the classification.

The intrinsic luminosity is shown in the X-ray light curve (Fig.
\ref{fig:sxray_lightcurve}) and it is sub-luminous compared to the majority of
long GRBs. The inset in Fig. \ref{fig:sxray_lightcurve} shows the luminosity
distribution at 10 ks. The sub-samples comprise of 333 lGRBs, 19 sGRBs, and
GRB~111117A. The sample of lGRBs are from \citet{Evans2007, Evans2009} and the
sample of sGRBs are compiled from \citet{Kann2011, Berger2014} and
\citet{DAvanzo2014a}. The mean and the 1-$\sigma$ dispersions of the samples are
log(L$_{lGRB}$)$ = 46.59 \pm 0.87$ erg s$^{-1}$ and log(L$_{sGRB}$)$  = 44.96
\pm 0.94$ erg s$^{-1}$ for the lGRB and sGRB samples respectively. GRB~111117A
had a log(L) $= 44.95$ erg s$^{-1}$ at 10 ks which is very close to the peak of
the sGRB luminosity distribution at 10 ks, but an outlier from the lGRG
distribution, further supporting the short classification.

\citet{Bromberg2013} investigated the degree to which the long and short
population distributions overlap and quantified the certainty in class
membership. According to \citet{Bromberg2013}, GRB~111117A has $96_{-5}^{+3}$
percent probability of being a sGRB. Compared to the other two sGRB candidates
at high redshift, GRB~060121 ($T_{90}=1.97 \pm 0.06$~s)
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB~090426 ($T_{90}=1.28 \pm 0.09$~s)
\citep{Antonelli2009, Levesque2010, Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$
per cent), the certainty in class membership for GRB~111117A is much higher.

Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
intermediate duration bursts. This comes from both events having very soft
spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
bursts are very clearly related in their properties to lGRBs
\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
mergers. GRB~111117A is also securely classified as a sGRB according to the
\citet{Horvath2010} classification scheme.

@@

The large offset of GRB~111117A relative to the host center derived in
\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
galaxy-scale gas providing the X-ray absorption. Along with the low dust content
of the host, the large offset from the host center indicates that the high $N_H$
arises because the density in the GRB surroundings is high, or because the light
from the afterglow traverses a region of dense gas. Assuming a dust-to-metals
ratio typical for the Milky Way, the X-ray derived $N_H$ corresponds to
significant extinction along the afterglow sight line ($A_V > 10$)
\citep{Predehl1995}, which is contrasted to the absence of dust found for the
SED fit and supported by the detection of \lya{}. The discrepancy between the extinction derived from the GRB afterglow
and the one obtained using galaxy-wide measures is also seen in the range of
ratios found for lGRBs \citep{Perley2013a, Covino2013}.

@@

The simultaneus detection of \lya{} and \ha{} allows us to put constraints of
the escape fraction of \lya{}, $f_{esc}$(\lya). Using the intrinsic ratio
between \ha{} and \lya{}, assuming case B recombination
\citep{Brocklehurst1971}, and the measured fluxes from the spectrum, we find
$f_{esc}$(\lya)$ < 0.06$. While the $f_{esc}$(\lya) scales with the dust column
\citep{Hayes2011}, the resonant scattering of \lya{}-photons with neutral
hydrogen makes the effective path length of \lya{} is longer than \ha{}
\citep{Atek2009}. This makes $f_{esc}$(\lya) an unreliable proxy for dust
column, especially at low dust columns \citep{Atek2014} where the geometry and
dynamics of the \hi{} within the galaxy will affect the \lya{} path the most.
The $f_{esc}$(\lya) inferred for the host is entirely consistent with what is
found for field galaxies with similar properties \citep{Oyarzun2017}. The same
authors also find that \lya{} emitting galaxies have mostly little dust,
consistent with what inferred from the SED-fit, see Sect. \ref{SED}. The
centroid of the Ly$\alpha$ emission is found to be redshifted by
$\sim$~220~km~s$^{-1}$ with respect to systemic, which is similar to what is
found for long GRB hosts \citep{Milvang-Jensen2012a} and Lyman Break Galaxies
\citep{Shapley2003a} in which the outflow is attributed to star formation.

@@

The redshift distributions of GRBs are powerful informants on both the
conditions which drive the formation of these events, but also on the potential
influence these cosmic explosions have on the evolution of the universe. Due to
elevated brightness of lGRBs compared to sGRBs \citep{Berger2014} and their
tendency to be associated with the most star-forming and therefore dense regions
in their hosts \citep{Fruchter2006, Lyman2017}, the afterglows of lGRBs are
easier to observe than their sGRB counterparts where only a single burst have a
redshift measurement from the GRB afterglow \citep{Cucchiara2013,
	DeUgartePostigo2014c}. Correspondingly, the redshift distribution of lGRBs is
better determined than that of sGRBs \citep[e.g., see][]{Jakobsson2012a,
	DAvanzo2015, Perley2016d}.

@@

Using the complete, flux-limited selection of \textit{Swift}-detected bursts
from \citet{DAvanzo2014a}, excluding GRB~111117A and the likely non-sGRB
GRB~090426, we have a statistically homogeneous sample from which we can address
the implications of the redshift of GRB~111117A. This sample includes sGRB
originating in star-forming galaxies, elliptical galaxies, and apparent hostless
sGRBs. Out of the 14 hosts in the sample, 10 (71 per cent) have measured
redshifts and $L_B/L^{\star}_{B}$. Of the complete sample, the host of
GRB~111117A  is brighter than 80 per cent of the hosts with measured
$L_B/L^{\star}_{B}$. Even if we conservatively assume that \textit{all} the
hosts missing $L_B/L^{\star}_{B}$ are brighter than the host of GRB~111117A, the
host is still brighter than > 60 per cent of sGRB hosts. For all 26 hosts with
$L_B/L^{\star}_{B}$ from \citet{Berger2014}, the host of GRB~111117A is brighter
than 73 per cent.

@@

A fraction of the bursts missing redshift are host-less but appear to be
spatially correlated with galaxies that are likely at moderate redshifts
\citep{Tunnicliffe2014}, but should some of the remainder be at high redshift,
the missed fraction will increase. If we assume that \textit{all} the bursts
that are missing redshifts are at high-$z$ and missed due to host faintness, 10
out of 14 hosts in the complete sample (71 per cent) would be missed at $z =
2.211$. This serves as an upper limit on the fraction of missed burst redshifts
at high-$z$. Conversely, if all bursts missing redshift are at low redshift and
missed for other reasons, 6 out of 14 hosts (43 per cent) would be missed at $z
= 2.211$. The two limits indicate that we would miss between 43 and 71 per cent
of \textit{Swift}-detected sGRB hosts at $z \sim 2$ due to host faintness.
